---
output:
  html_notebook: default
  html_document: default
---
Project Report 

Background
We are given a data set that is related to the red and white variants
of the "Vinho Verde" wine from Portugal. In the dataset, the data
set contains the various chemical test that were done on the wines
as well as an overall assesment on from at least three professional 
taste testers. 

Initial EDA
```{r}
red <- read.csv('winequality-red.csv', sep=";")
white <- read.csv('winequality-white.csv', sep=";")
red2 <- read.csv('winequality-red.csv', sep=";")
white2 <-read.csv('winequality-white.csv', sep=";")
#Adding a column for color since we're comparing quality

red['color'] <- 'red'
white['color'] <- 'white'
redandwhite <- rbind(red,white)
```

First we read the two .csv files for the red and white wine data, while adding a new column into each file called 'color' to help us distinguish what is what. 

```{r}
names(redandwhite)
```

Let's check and see if there are any missing values before diving deeper.
```{r}
sum(is.na(redandwhite))
```

We don't have any null values, so now I'm going to dig deeper by looking at the values. 

```{r}
summary(redandwhite)
```


```{r}
str(redandwhite)
```

Some Observerations so far from the summary: 
-There are no basic wines in this dataset (Basic in pH usually means anything greater than 7 on the pH scale)
-The wine is pretty acidic, the range is 2.7-4, and the mean is 3.2
-Free SO2 (Sulfur Dioxide) has an outlier of 289.00 ppm. This is very high since the 3rd quarter is only at 41.00 and the mean is at 30.53
-There is a very sweet wine at 65.800 in residual sugar, the 75% residual sugar is at 8.1
-The alcohol levels are pretty close to each other, the one with the lowest amount of alcohol is at 8 and it ranges all the way to 14.
-Density is pretty consistent to each other. The min is at 0.9871 and the max being at 1.03.
-Quality ranges from 3 to 9. 

Initial EDA. Next I'll be taking a deeper look of the qualities that were measured. 

```{r}
RedQuality <- summary(red$quality)
RedQuality
```

```{r}
ggplot(redandwhite, aes(x=quality, fill = color)) + geom_histogram(alpha =0.5, position = "identity", binwidth = 1) + scale_x_continuous(breaks = seq(3,9,1), lim =c(3,10))
```

We know that there a different number of obserations of red and white wine in the dataset. But just by looking at this, you can tell that it is at a normal distribution. 

Checking out the alcohol level: 

```{r}

summary(redandwhite$alcohol)
```

```{r}
ggplot(redandwhite, aes(x=alcohol, fill = color)) + geom_histogram(alpha = 0.5, position = "identity", binwidth = 1) + scale_x_continuous(breaks = seq(8,15,1), lim= c(8,15))
```


Looking at the numbers for density, I have a feeling that if I were to model this, the graphs would look weird. 

```{r}
ggplot(redandwhite, aes(x=density, fill = color)) + 
  geom_histogram(alpha =0.5, position = "identity", binwidth = 0.0002)+
  scale_x_log10(lim = c(min(redandwhite$density), 1.00370),
                breaks = seq(min(redandwhite$density), 1.00370, 0.005))

```
When we look at the summary for the density, there are outliers. Initially I graphed this as a regular histogram but it was not clear enough for me to see the trends. After searching, I decided to use log10. Red wine's density normally distributed while the density distribution for white wine is bimodal with two maximum values. 

Taking a look at Sulfur Dioxide

```{r}
summary(redandwhite$free.sulfur.dioxide)
```


As mentioned previously, looking at the Sulfur Dioxide, we have a very large outlier of 289.00 as the max. 

```{r}
ggplot(redandwhite, aes(x =free.sulfur.dioxide, fill = color)) +
  geom_histogram(alpha = 0.5, position = "identity", binwidth =1) +
  scale_x_continuous(breaks = seq(min(redandwhite$free.sulfur.dioxide),
                                  max(redandwhite$free.sulfur.dioxide), 20))
```

THis is just free sulfur dioxide, we should also take a look at the total amount. 

```{r}
summary(redandwhite$total.sulfur.dioxide)
```

```{r}
ggplot(redandwhite, aes(x = total.sulfur.dioxide, fill = color))+ geom_histogram(alpha = 0.5, position = "identity", binwidth = 1) + scale_x_continuous(breaks = seq(min(redandwhite$total.sulfur.dioxide), max(redandwhite$total.sulfur.dioxide),20))
```
Breaking down both the free and the total Sulfur Dioxide, there is a normal distribution in Sulfur Dioxide for White wine, whereas with the red wine has a log-normal distribution. While it seems like the red wine total sulfur dioxide has no outliers, looking deeper for the red wine there is a total amount of 290ppm. 

Looking at Correlation between factors in Red and White Wine

```{r}
redtitle <- "Red Wine Correlation"
corrplot(redmat, order = "AOE", method = "color", addCoef.col ='black', number.cex = 0.5, title = redtitle, line = -1)
```
Looking at this for red wine we can infer that:
-Total Sulfur Dioxide and Free Sulfur Dioxide that there is a relationship between them. 
-pH and fixed acidity are related 
-Alcohol and quality have an inverse relationship
```{r}
whitetitle <- "White Wine Correlation"
corrplot(whitemat, order = "AOE", method = "color", addCoef.col ='black', number.cex = 0.5, title = whitetitle, line = -1)
```
-Low Density means high alcohol content
-High density means high residual sugar 
-Again, there is a position relationship with the free and total sulfur dioxide. 

Creating Boxplots and 
checking out Outliers

