---
title: "R Notebook"
output: html_notebook
---

Project Background: 

Wine has been around for a long time! It is a drink that most people come to enjoy while pairing it with a good steak or just something to enjoy while watching some Netflix on a Wednesday night. Wine is made through the process of fermentation with the sugars in any type of fruit juice. While wine can be made with any type of fruit, grapes are mostly used due to the amount of fermentable sugar grapes have.

The food and beverage industry is a huge industry anywhere. The wine market is a huge industry and it is estimated to be about 39.8 billion dollars in the US. Thanks to the power and popularity of data science and its applications, a company can make use of the data they find to stay ahead of their competition or even to automate some of the smaller processes in the product development process. 

This is a small project of how data science can aid the wine business. 

About the dataset: 

The two datasets contain different characteristics which are the physico--chemical and sensorical of red and white wines. The product is called Vinho Verde, the data can be found in the UCI Machine Learning Repository. 

```{r}
#Importing all the required libraries
#ggplot2 -> for modeling purposes
#caret -> ML purposes for the model
#plsRglm -> for ML purposes for the model
#corrplot -> heat matrix 
library(ggplot2)
library(caret)
library(tidyverse)
library(plsRglm)
library(corrplot)
```


```{r}
#Reading the datasets
#Stored into two different variables
#One will be combined later on 

red <- read.csv('winequality-red.csv', sep=";")
white <- read.csv('winequality-white.csv', sep=";")
red2 <- read.csv('winequality-red.csv', sep=";")
white2 <-read.csv('winequality-white.csv', sep=";")
#Adding a column for color since we're comparing quality

red['color'] <- 'red'
white['color'] <- 'white'

#Combining both the datasets 
redandwhite <- rbind(red,white)
```

Now that we have our data loaded in let's take a glimpse at it. 

```{r}
#Just checking the combined data set instead of the individual data sets for now
dim(redandwhite)
names(redandwhite)
summary(redandwhite)
```

```{r}
str(redandwhite)
```

Some Observerations so far from the summary: 
-There are no basic wines in this dataset (Basic in pH usually means anything greater than 7 on the pH scale)
-The wine is pretty acidic, the range is 2.7-4, and the mean is 3.2
-Free SO2 (Sulfur Dioxide) has an outlier of 289.00 ppm. This is very high since the 3rd quarter is only at 41.00 and the mean is at 30.53
-There is a very sweet wine at 65.800 in residual sugar, the 75% residual sugar is at 8.1
-The alcohol levels are pretty close to each other, the one with the lowest amount of alcohol is at 8 and it ranges all the way to 14.
-Density is pretty consistent to each other. The min is at 0.9871 and the max being at 1.03.
-Quality ranges from 3 to 9. 

Understanding the variables that we're working with. To start with the initial data analysis, I'll be making a corrplot showing the relationship between the variables that were tested. 

```{r}
redmat <- round(cor(red2),2)
redtitle <- "Red Wine Correlation"
corrplot(redmat, order = "AOE", method = "color", addCoef.col ='black', number.cex = 0.5, title = redtitle, line = -1)
```

```{r}
whitemat <-round(cor(white2),2)
whitetitle <- "White Wine Correlation"
corrplot(whitemat, order = "AOE", method = "color", addCoef.col ='black', number.cex = 0.5, title = whitetitle, line = -1)
```

Some observation looking at this: 
For Red Wine:
-Density and Residual Sugar are positive with a 0.36 -> we can infer that the density is affected by the residual sugar
-pH and fixed acidity have an inverse relationship.
-

```{r}
ggplot(redandwhite, aes(x=quality, fill = color)) + geom_histogram(alpha =0.5, position = "identity", binwidth = 1) + scale_x_continuous(breaks = seq(3,9,1), lim =c(3,10)) + ggtitle("Quality Scores among Red and White Wine")
```

```{r}
ggplot(redandwhite, aes(x=alcohol, fill = color)) + geom_histogram(alpha = 0.5, position = "identity", binwidth = 1) + scale_x_continuous(breaks = seq(8,15,1), lim= c(8,15)) + ggtitle("Alcohol Levels among Red and White Wine")
```

```{r}
ggplot(redandwhite, aes(x=density, fill = color)) + 
  geom_histogram(alpha =0.5, position = "identity", binwidth = 0.0002)+
  scale_x_log10(lim = c(min(redandwhite$density), 1.00370),
                breaks = seq(min(redandwhite$density), 1.00370, 0.005)) + ggtitle("Density Levels Among Red and White Wine")
```

```{r}
ggplot(redandwhite, aes(x =free.sulfur.dioxide, fill = color)) +
  geom_histogram(alpha = 0.5, position = "identity", binwidth =1) +
  scale_x_continuous(breaks = seq(min(redandwhite$free.sulfur.dioxide),
                                  max(redandwhite$free.sulfur.dioxide), 20))+ ggtitle("Red and White Wine Free SO2 levels")
```

```{r}
ggplot(redandwhite, aes(x = total.sulfur.dioxide, fill = color))+ geom_histogram(alpha = 0.5, position = "identity", binwidth = 1) + scale_x_continuous(breaks = seq(min(redandwhite$total.sulfur.dioxide), max(redandwhite$total.sulfur.dioxide),20)) + ggtitle("Total SO2 in Red and White Wine")
```

